### 2014-04-02 paper/discussion.tex
comments on latter half of the paper

The promise of 3D printing is that it allows for the manufacture of geometries that are impossible to create using traditional techniques like injection molding.  Pipes leverage that promise: designing the internal structure of an object was not previously feasible.  As printers at the hobbyist and corporate levels become more sophisticated and able to fabricate intricate geometries of ever more materials, we believe that our tool will still assist in design tasks.  Even if printers are capable of laying down conductive, optically transparent, or luminescent materials in-place, the process of designing their paths is still crucial.

By crossing pipe types, topologies, and inserted media, we create a space of possible inputs and outputs that can be pipe-mediated.  We arrange our I/O discussion around the modes of input or output (the five senses, touch/tapping/pressure/etc.).  We also suggest new points in the design space of pipe-mediated interaction that remain unexplored.

@@

We describe potential pipe-based outputs organized by the five senses: visual, aural, haptic, and olfactory/gustatory.  In addition to outputs we mention explicitly, any outputs possible using traditional electronic components are possible using conductive material in pipes.

@@

\emph{Visual} outputs can be mediated by gases, liquids, solid, particulates, or threadables.  Pneumatic actuation (as in PneUIs \cite{Yao-pneui} or Harrison's latex buttons \cite{Harrison-buttons}) provides output, but gases carrying smoke (as in the Aireal \cite{Sodhi-aireal} demo) also offers visual feedback.  Splitting and mixing of colored liquids can be used as a display tactic; additionally mechanical motion can be fluid-mediated.  Liquids can also be mixed in specific concentrations by servo-controlled gate structures as in Figure \ref{fig:direct}.  Printed optics \cite{Willis-printedoptics} explored visual output at the ends of pipes, and  EL wire can be threaded through custom paths to create neon signs, as in our neon robot (Figure \ref{fig:neon}).

@@

>>>>

The promise of 3D printing is that it allows for the manufacture of geometries that are impossible to create using traditional techniques like injection molding.  Pipes leverage that promise: designing the internal structure of an object was not previously feasible. \bjoern{overly broad claim. companies are designing the "interior " when they make complex shelled parts using injection molding. But that process has lots of limitations as to the geometries that are possible --- there's a whole industry of specialists concerned with ``design for manufacturability'' (DFM).} As printers at the hobbyist and corporate levels become more sophisticated and able to fabricate intricate geometries of ever more materials, we believe that our tool will still assist in design tasks. \bjoern{oblique, and could be phrased more positively.} Even if printers are capable of laying down conductive, optically transparent, or luminescent materials in-place, the process of designing their paths is still crucial.

By crossing pipe types, topologies, and inserted media, we create a space of possible inputs and outputs that can be pipe-mediated.  We arrange our I/O discussion  around the modes of input or output (the five senses, touch/tapping/pressure/etc.).  We also suggest new points in the design space of pipe-mediated interaction that remain unexplored. \bjoern{need to motivate first why the discussion is going to focus on inputs and outputs - why is this the logical topic to return to? Maybe something like this: ``By crossing pipe types, topologies, and inserted media, many pipe-mediated interactions emerge. We demonstrated a few such possibilities by designing and fabricating example objects with the help of our pipe routing design tool. However, there are many additional possibilities for leveraging sensing and actuation techniques from the HCI literature and extending their use in 3D-printed objects. We briefly survey such possibilities for output and input.''}

@@

Pipe-based outputs can address the five senses: visual, aural, haptic, and olfactory/gustatory.  In addition to outputs we mention explicitly, any outputs possible using traditional electronic components are possible using conductive material in pipes.

@@

\emph{Visual} outputs can be mediated by gases, liquids, solid, particulates, or threadables.  Pneumatic actuation (as in PneUIs \cite{Yao-pneui} or Harrison's latex buttons \cite{Harrison-buttons}) provides output, but gases carrying smoke (as in the Aireal \cite{Sodhi-aireal} demo) also offers visual feedback \bjoern{mischaracterization of aireal - they only added smoke to show what a vortex looks like; the technique is all about invisible, haptic sensation}.  Splitting and mixing of colored liquids can be used as a display tactic; additionally mechanical motion can be fluid-mediated.  Liquids can also be mixed in specific concentrations by servo-controlled gate structures as in Figure \ref{fig:direct}.  Printed optics \cite{Willis-printedoptics} explored visual output at the ends of pipes, and  EL wire can be threaded through custom paths to create neon signs, as in our neon robot (Figure \ref{fig:neon}) \bjoern{are we going to repeat things here we already talked about? Maybe we should just focus on ideas one could build that we haven't built yet}.

@@

\bjoern{We might pull back on the scope of this section a bit and think about which directions here are most closely related to pipes and most likely successful;  and which are more tangential or speculative - for example, Aireal is about feedback in free space and we are all about designing tangible objects; so maybe that's a candidate for cutting.}
